---
title: "Probabilistic ensemble forecasting of Australian COVID-19 cases"
author: "Rob J Hyndman"
date: ISF 2021
fontsize: 14pt
classoption: aspectratio=169
toc: false
output:
  binb::monash:
    fig_height: 4.33
    fig_width: 7
    colortheme: monashwhite
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 88)
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
localcases <- readRDS("localcases.rds") %>%
  select(date, state, n) %>%
  filter(state != "AUS")
```

## Australian Health Protection Principal Committee

\begin{block}{}The \textbf{Australian Health Protection Principal Committee} is the key decision-making committee for national health emergencies. It comprises all state and territory Chief Health Officers and is chaired by the Australian Chief Medical Officer.
\end{block}

\begin{alertblock}{COVID-19 forecasting group}
\begin{multicols}{3}\small
\begin{itemize}\tightlist
\item Peter Dawson
\item Nick Golding
\item Rob J Hyndman
\item Dennis Liu
\item James M McCaw
\item Jodie McVernon
\item Pablo \rlap{Montero-Manso}
\item Robert Moss
\item Mitchell \rlap{O'Hara-Wild}
\item David J Price
\item Joshua V Ross
\item Gerry Ryan
\item Freya M Shearer
\item Tobin South
\item Ruarai Tobin
\end{itemize}
\end{multicols}\vspace*{-0.2cm}
\end{alertblock}

## Data sources

* Case-level data of all positive COVID-19 tests: onset and detection times.
* Daily population mobility data from Google, Apple & Facebook
* Weekly non-household contact surveys
* Weekly behavioural surveys
* Daily case data from ?? countries via the Johns Hopkins COVID-19 repository

## Case numbers
\fontsize{12}{12}\sf

```{r, echo=TRUE}
localcases %>% filter(state=="VIC", date >= "2020-07-01")
```

## Case numbers

```{r, echo=TRUE, fig.height=3}
localcases %>% autoplot(n)
```

## Mobility data

\full{mobility}

## Mobility data

\full{google}

## Mobility data

\full{facebook}

## Survey data

\full{macrodistancing}

## Survey data

\full{microdistancing}

## Blah

We use case-level data of all Australian positive COVID cases, along with nationwide surveys and mobility data from Google, Facebook and Apple. Three separate models have been built: (1) a stochastic susceptible-exposed-infectious-recovered (SEEIIR) compartmental model; (2) a stochastic epidemic model; and (3) a global autoregressive model based on public case data from 31 countries. These are then combined into a mixture ensemble to generate probabilistic forecasts of daily cases which are provided to the Australian governments each week. I will discuss the ensemble forecasting aspects of this work and how we evaluate the results.



## Thanks!

\placefig{10}{1.4}{trim = 30 0 0 0, clip=TRUE, width=2cm, height=2.5cm}{mitch}

## More information
\fontsize{18}{20}\sf

 * Slides and papers: **robjhyndman.com**
 * Packages: **tidyverts.org**
 * Forecasting textbook using fable package: **OTexts.com/fpp3**

\begin{textblock}{8}(7.6,4.8)
\begin{alertblock}{Find me at ...}
\href{https://twitter.com/robjhyndman}{\faicon{twitter} @robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github}  @robjhyndman}

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope}  rob.hyndman@monash.edu}
\end{alertblock}
\end{textblock}
\vspace*{10cm}
